Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-12-29 demystify.tex
updates

\section{GREs Are Optional and Do Not Matter} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it. Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it. Being optional really means optional, and not taking it will not hurt your application.

However, if you took it and have really good scores then it might be worth it to include and talk about them in your application, but don't expect them to make much difference (many adcom reviewers never took it). But if your scores are bad, then you should not include them in your application, which can be a \red{red flag}.

@@

Unless your degrees are from the US or certain countries such as \href{https://www.gmu.edu/international/english-language-requirements}{these}, you will need to take standardized English tests. On one hand you will need to show some level of English proficiency, but on the other hand, you do not need to have very good scores in these tests (many adcom members themselves were once international students and struggled with English).
You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}).
Just as with grades (\autoref{chap:your-school}) and GRE, having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.

@@

While being ambitious and aiming high is good, you should understand how PhD admission works (e.g., by reading \emph{this} handbook and realizing things such as having a good GPA or GRE doesn't mean much to top programs), that in the US there are many good schools, and just be realistic.

@@

\paragraph{School \autoref{chap:your-school}} Adcom already knows about your school, which is a plus (see \autoref{chap:your-school}). You are also more familiar with the US education system and academic culture (\autoref{sec:cultural}).

@@

\paragraph{Test Scores} GREs are often not required for MS applications.  However, some schools might still require them, and you should check with the school you're applying to.  If you have a low GPA, a high GRE score can help offset that. As with PhD, English proficiency tests (TOEFL/IELTS) are required for international students (but just passing the mimimum requirement is often enough \autoref{sec:english-tests}). Note that just as with PhD, Englih tests are are waived if you did your undergrad at a US university.

>>>>

\section{GREs Are Optional and Do Not Matter}\label{sec:gre}

While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it. The reason is that GRE scores do not correlate well with research ability, which is the most important factor for PhD admission. It is also worth noting that many faculty members themselves did not take the GRE or have bad scores.

Thus, if you have bad GRE scores or haven't taken the GRE, then don't waste time (re)taking it. Being optional really means optional, and not taking it will not hurt your application.
However, if you took it and have really good scores then it might be worth it to include (and perhaps talk about) them in your application, but don't expect them to make much difference. But if your scores are bad, then you should not include them in your application, which can be a \red{red flag}.

@@

Unless your degrees are from the US or certain countries such as \href{https://www.gmu.edu/international/english-language-requirements}{these}, you will need to take standardized English tests. On one hand you will need to show some level of English proficiency, but on the other hand, you do not need to have very good scores in these tests (many adcom members themselves were once international students and struggled with English).
You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}).
Just as with grades (\autoref{chap:your-school}) and GRE (\autoref{sec:gre}), having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.

@@

While being ambitious and aiming high is good, you should understand how PhD admission works (e.g., by reading \emph{this} handbook and realizing things such as having a good GPA (\autoref{sec:gpa}) or GRE (\autoref{sec:gre}) doesn't mean much to top programs), that in the US there are many good schools, and just be realistic.

@@

\paragraph{School \autoref{chap:your-school}} Adcom already knows about your school, which is a plus. You are also more familiar with the US education system and academic culture (\autoref{sec:cultural}).

@@

\paragraph{Test Scores} Unlike PhDs, which often do not require GREs, MS programs often require GREs.  However, some schools do make them optional, and you should check with the school you're applying to.  If you have a low GPA, a high GRE score can help offset that.

As with PhD, English proficiency tests (TOEFL/IELTS) are required for international students (but just passing the minimum requirement is often enough \autoref{sec:english-tests}). Note that just as with PhD, English tests are are waived if you did your undergrad at a US university.
